Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-18 forecast/paper/paper.tex
FINISH discussion

Along the lines of regularization, it is also worth noting that the use of
$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, 
\eqref{eq:hotspot_ar_x} did not improve their generally performance at all
(experiments not shown).  This is likely due to the fact that the number of
training samples is quite large compared to the number of features (6,426
training samples and only 3--6 features).

@@

\subsection{Performance in Up, Down, and Flat Periods}

@@

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/upswing-histogram-1.pdf}
  \caption{Histogram of the difference in WIS for the AR model and that for each
    indicator model, stratified by up, down, or flat periods, measured in terms
    of case activity. (Larger values are better for the indicator model.) The
    y-axis is on the log scale to emphasize tail behavior.} 
  \label{fig:up_down_flat}
\end{figure}

@@

As described in the methods section, each of the indicators we examine could be
said to measure aspects on disease progression that would precede a positive
test. That is, we imagine that these signals should ``lead'' cases. It is
entirely reasonable to imagine that, prior to an increase of confirmed COVID-19
tests reported by public health in a particular location, we would see an
increase in insurance claims for COVID-19-related medical visits.  However, it
may well be the case that this leading behavior is different during different
periods. In fact, we find that the ``leadingness'' of an indicator (degree to
which it leads) tends to be more pronounced during down and flat periods than up
periods, a plausible explanation for the decreased performance noted above.

In order to measure the leadingness on af indicator, we examine its cross
correlations to case rates over a short time window

 the cross
correlation function. Essentially this examines the Pearson correlation between

shifted version of the indicator and the case rate. So if $X_{\ell,t}$ is
shifted forward by 7 days, and the magnitude of the correlation of that shifted series with
$Y_{\ell,t}$ (unshifted) is high, we would say that $X_{\ell}$ leads $Y_{\ell}$
by 7 days. This process is repeated for a range of shifts, forward and backward,
resulting in leadingness and laggingness scores. Further details are given in
the supplement.

Figure \ref{fig:leading-lagging} shows the correlations between the difference
in the WIS of the AR model relative to $F$ with the leading and laggingness
scores stratified by whether the target is classified as up, down, or flat. One
would expect that increases in $\WIS(AR) - \WIS(AR)$ would be positively
correlated with leadingness. This relationship is strongest during down periods
and weakest during up periods with flat in the middle. In fact, for each
indicator, the strength of correlations between improved performance and
leadingness can be ordered down $>$ flat $>$ up. The relationship with
laggingness is the opposite, though less pronounced. CHNG-CLI's forecast
performance doesn't correlate very strongly with laggingness while any
laggingness in Google-AA during down and flat periods results in realtively
worse forecasting performance.

Because a time series can, to a degree, both simultaneously leading and lagging,
we also examine the correlation between performance and the difference between
leadingness and laggingness. Figure \ref{fig:diff-btw-leading-lagging} shows
this behavior. Again, indicators that are more leading then lagging tend to
are positively correlated with improved relative performance. This relationship
is most pronounced during periods of decreasing case rates and attenuates in
other scenarios.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/leading-and-lagging-1.pdf}
  \caption{Correlation of the difference in WIS with the ``leadingness'' and
    ``laggingness'' of the indicator at the target date.}
  \label{fig:leading-lagging}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/diff-in-lead-lag-1.pdf}
  \caption{Correlation of the difference in WIS with difference between the
    ``leadingness'' and ``laggingness'' of the indicator at the target date.}
  \label{fig:diff-btw-leading-lagging}
\end{figure}

@@

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the  
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and Internet search trends) is undoubtedly ``yes''. However, there are  
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated appear to be the ``silver bullet'' that 
one might have hoped for, revolutionizing the tractability of the prediction 
problem, rendering it easy when it was once hard (in the absence of auxiliary 
information).  Rather, the gains in accuracy from the indicator models (over an
autoregressive model based only on past case rates) appear to be nontrivial, and
consistent across modes of analysis, but modest.  In forecasting, the indicator 
models appear to be most useful in periods in which case rates are flat or
trending down, rather than periods in which case rates are trending up (as one 
might hope to see is the benefit provided by a hypothetical ``leading
indicator'').

As mentioned previously, the indicator models could likely be improved by using 
location-specific effects, as well as using nowcasting techniques to estimate
finalized indicator values before using them as features (in order to account
for backfill in the claims-based signals in particular).  Beyond this, it is
certainly possible that more sophisticated models for forecasting or hotspot
prediction would lead to different results, and even different
insights. Natural directions to explore include using interactions, using
multiple indicators in a single model, and leveraging HRR demographics or
mobility patterns.  That said, we are doubtful that more sophisticated modeling
techniques would change the ``topline'' conclusion---that auxiliary indicators 
can provide nontrivial, consistent, but modest gains in forecasting and
hotspot prediction. Whether a more sophisticated model would be able to leverage
the indicators in such a way as to change some of the finer conclusions---e.g.,
by offering clear improvements in periods in which cases are trending up---is 
less clear to us.

Reiterate importance of AS OF

Not the end of the story.  Should we even be forecasting case rates?  More
careful models---combine hotspots and forecasting?

[  Finish by reiterating the importance of vintage data.  Emphasizing the rigor
or our approach, how we have provided a template for others to be able to do the
same thing with new models ]

% RJT: some text commented out below.  May be useful for finishing off the
% discussion

% This paper is as much about demonstrating how one might go about answering
% such questions about feature importance in forecasting and hotspot prediction 
% general, as providing an empirically rigorous answer to this precise question
% in the context of the COVID-19 pandemic.

% We conclude by observing that the approach we have taken in this paper is
% somewhat opposite to that of much of the COVID-19 forecasting literature.  We
% have chosen to consider only very simple forecasting models while devoting
% most of our effort to accounting for as much of the complexity of the data and 
% evaluation as possible.  By contrast, many papers focus on very complicated
% forecasting approaches but then evaluate them under unrealistic, retrospective 
% conditions.

>>>>

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, 
\eqref{eq:hotspot_ar_x} did not improve their generally performance at all
(experiments not shown).  This is likely due to the fact that the number of
training samples is large compared to the number of features (6,426 training
samples and only 3--6 features).

@@

\subsection{Evaluation in Up, Down, and Flat Periods}

@@

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/upswing-histogram-1.pdf}
  \caption{Histogram of the difference in WIS for the AR model and that for each
    indicator model, stratified by up, down, or flat period, measured in terms
    of case trends. Note that \textit{larger} differences here are better for
    each indicator model.  The y-axis is on the log scale to emphasize tail 
    behavior.}   
  \label{fig:up_down_flat}
\end{figure}

@@

As described in the methods section, each of the indicators we examine could be
said to measure aspects on disease progression that would precede a positive
test. That is, we imagine that these signals should ``lead'' cases. It is
entirely reasonable to imagine that, prior to an increase of confirmed COVID-19
tests reported by a public health authority in a particular location, we would
see an increase in medical insurance claims for COVID-related outpatient
visits. However, it may well be the case that such behavior is different
during different periods. In fact, we find empirically that the ``leadingness''
of an indicator (degree to which it leads case activity) tends to be more
pronounced in down or flat periods than up periods, a plausible explanation
for the decreased performance in up periods noted above.

In the supplement, we show how to define a quantitative score to measure the
leadingness of an indicator, at any time $t$ and any location $\ell$, based on
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity.  Figure 
\ref{fig:leading} displays correlations between the leadingness score of an 
indicator and the WIS difference (AR model minus an indicator model), stratified
by whether the target is classified as up, down, or flat.  One would expect that  
the WIS difference would be positively correlated with leadingness.  Somewhat
surprisingly, this relationship turns out to be the strongest in down periods,
and weakest in up periods.  In fact, it is very nearly the case that for each
indicator, the strength of correlations only decreases as we move from down to
flat to up  periods.  In the supplement, we extend this analysis by examining
analogous ``laggingness'' scores, but we do not find as clear patterns.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/leading-and-lagging-1.pdf}
  \caption{Correlation of the difference in WIS with the ``leadingness'' of the
    indicator at the target date, stratified by up, down, or flat period.} 
  \label{fig:leading}
\end{figure}

@@

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and Internet search trends) is undoubtedly ``yes''. However, there are
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated appear to be the ``silver bullet'' that one
might have hoped for, revolutionizing the tractability of the prediction
problem, rendering it easy when it was once hard (in the absence of auxiliary
information).  Rather, the gains in accuracy from the indicator models (over an
autoregressive model based only on past case rates) appear to be nontrivial, and
consistent across modes of analysis, but modest.  In forecasting, the indicator
models are found to be most useful in periods in which case rates are flat or
trending down, rather than periods in which case rates are trending up (as one
might hope to see is the benefit provided by a hypothetical ``leading
indicator'').

As described previously, it is likely that we could improve the indicator models
by using location-specific effects, as well as using nowcasting techniques to
estimate finalized indicator values before we use them as features (to account
for backfill in the claims-based signals in particular).  Beyond this, it is
certainly possible that more sophisticated models for forecasting or hotspot
prediction would lead to different results, and possibly even different
insights. Natural directions to explore include using multiple indicators in a
single model, allowing for interaction terms, and leveraging HRR demographics or
mobility patterns.  That said, we are doubtful that more sophisticated modeling
techniques would change the ``topline'' conclusion---that auxiliary indicators
can provide nontrivial, consistent, but modest gains in forecasting and hotspot
prediction. Whether a more sophisticated model would be able to leverage the
indicators in such a way as to change some of the finer conclusions---e.g., by
offering clear improvements in periods in which cases are trending up---is less
clear to us.

We reiterate the importance of using vintage data for rigorous backtesting. Data
sources that are relevant to public health surveillance are often subject to
revision, sometimes regularly (such as medical claims data) and sometimes
unpredictably (such as COVID-19 case reports).  When analyzing models that are
designed to predict future events, if we train these models and make predictions
using finalized data, then we are missing a big part of the story.  Not only
will our sense of accuracy be unrealistic, but certain models may degrade by a
greater or lesser extent when they are forced to work with vintage data, so 
backtesting them with finalized data may lead us to make modeling decisions that
may not be relevant for true test-time performance.

In this paper, we have chosen to consider only very simple forecasting models,
while devoting most of our effort to accounting for as much of the complexity of
the data and evaluation as possible.  In fact, our paper is as much about
demonstrating how one might address questions about model comparisons and
evaluation in forecasting and hotspot prediction in general, as it is about
providing rigorous answers to such questions in the context of COVID-19 case
rates in particular.  We hope that others will leverage our framework, and build
on it, so that it can be used to guide work that advances the frontier of
predictive modeling for epidemics and pandemics.

### 2021-06-18 forecast/paper/unused.tex
FINISH discussion

>>>>

Essentially this examines the Pearson correlation between 
shifted version of the indicator and the case rate. So if $X_{\ell,t}$ is
shifted forward by 7 days, and the magnitude of the correlation of that shifted series with
$Y_{\ell,t}$ (unshifted) is high, we would say that $X_{\ell}$ leads $Y_{\ell}$
by 7 days. This process is repeated for a range of shifts, forward and backward,
resulting in leadingness and laggingness scores. Further details are given in
the supplement.

Because a time series can, to a degree, both simultaneously leading and lagging,
we also examine the correlation between performance and the difference between
leadingness and laggingness. Figure \ref{fig:diff-btw-leading-lagging} shows
this behavior. Again, indicators that are more leading then lagging tend to
are positively correlated with improved relative performance. This relationship
is most pronounced during periods of decreasing case rates and attenuates in
other scenarios.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/diff-in-lead-lag-1.pdf}
  \caption{Correlation of the difference in WIS with difference between the
    ``leadingness'' and ``laggingness'' of the indicator at the target date.}
  \label{fig:diff-btw-leading-lagging}
\end{figure}
